### 2017-08-30 paper.tex
:computer: Coupled LB-DEM model

\subsection{Coupled LB-DEM for fluid-grain interactions}
The Lattice Boltzmann approach has the advantage of accommodating large particle sizes and the interaction between the fluid and the moving particles can be modelled through relatively simple fluid - particle interface treatments. Further, employing the Discrete Element Method (DE) to account for the particle/particle interaction naturally leads to a combined LB - DEM solution procedure. The Eulerian nature of the Lattice Boltzmann formulation, together with the common explicit time step scheme of both the Lattice Boltzmann and the Discrete Element makes this coupling strategy an efficient numerical procedure for the simulation of particle-fluid systems~\citep{Cook2004}. In order to capture the actual physical behaviour of the fluid-particle system, the boundary condition between the fluid and the particle is modelled as a non-slip boundary condition, i.e. the fluid near the particle should have similar velocity as the particle boundary. The solid particles inside the fluid are represented by lattice nodes. The discrete nature of lattice will result in stepwise representation of the surfaces. Very small lattice spacing is adopted to obtain smoother boundaries.

>>>>

\section{Coupled LB-DEM model}
\subsection{General}
The Lattice Boltzmann approach has the advantage of accommodating large particle sizes and the interaction between the fluid and the moving particles can be modelled through relatively simple fluid - particle interface treatments. Further, employing the Discrete Element Method (DEM) to account for the particle/particle interaction naturally leads to a combined LB - DEM solution procedure. The Eulerian nature of the Lattice Boltzmann formulation, together with the common explicit time step scheme of both the Lattice Boltzmann and the Discrete Element, makes this coupling strategy an efficient numerical procedure for the simulation of particle-fluid systems~\citep{Cook2004}. The LB-DEM coupling system is a powerful fundamental research tool for investigating hydro-mechanical physics in porous media flow~\citep{Han2013}. To capture the actual physical behaviour of a fluid-particle system, the boundary condition between the fluid and the particle is modelled as a non-slip boundary condition, i.e. the fluid near the particle should have a similar velocity as the particle boundary. The solid particles inside the fluid are represented as solid lattice nodes. The discrete nature of lattice will result in stepwise representation of the surfaces~\citep{Kumar2015}. A very small lattice spacing is adopted to obtain smoother boundaries.

The smallest DEM grain in the system controls the size of the lattice. In the present study, a very fine resolution of $d_{min}/h = 10$ is adopted. That is, the smallest grain with a diameter dmin in the system is discretized into 100 lattice nodes ($10h \times 10h$). This provides a very accurate representation of the interaction between the solid and the fluid nodes.

When combining the Discrete Element modelling of grain interactions with the lattice Boltzmann formulation, an issue arises. That is, there are now two time steps: $\Delta t$ for the fluid flow and $\Delta t_D$ for the particle movements. Since $\Delta t_D$ is normally smaller than $\Delta t$, $\Delta t_D$ is slightly reduced to a new value $\Delta t_s$ so that $\Delta t$ and $\Delta t_s$ have an integer ratio $n_s$:

\begin{align}
  \Delta t_s = & \Delta t / n_s \\
  n_s  = & [\Delta t/\Delta t_D]+1
\end{align}

This results in a subcycling time integration for the Discrete Element part.  $\Delta t$ every step of the fluid computation, $n_s$ sub-steps of integration are performed for DEM using the time step  $\Delta t_s$. The hydrodynamic force is unchanged during this sub-cycling.
